---
title: "Wine Magzine Project - Data Preprocessing - Matthew Lipinski"
output: html_notebook
---
Import initial dataset for preprocessing.
```{r}
library(readr)
winemag_data_first150k <- read_csv("c:/Users/mlipinsk/Documents/UST/SEIS 734-01 Data Mining & Pred. Analytics/Project/winemag-data_first150k.csv")
View(winemag_data_first150k)
```
Remove the record count column.
```{r}
WMDF9 <- winemag_data_first150k[, 2:11]
```
Dataframe for duplicates.  Returns 53,081 duplicate values.
```{r}
dups <- WMDF9[duplicated(WMDF9), ]
```
Remove duplicates.
```{r}
WMDF <- WMDF9[!duplicated(WMDF9), ]
```
Clean description.
```{r}
description2 <- stringi::stri_trans_general(WMDF$description, "latin-ascii")
```
New description.
```{r}
description3 <- as.data.frame(description2)
```
Clean designation.
```{r}
designation2 <- stringi::stri_trans_general(WMDF$designation, "latin-ascii")
```
New designation.
```{r}
designation3 <- as.data.frame(designation2)
```
Clean province.
```{r}
province2 <- stringi::stri_trans_general(WMDF$province, "latin-ascii")
```
New province.
```{r}
province3 <- as.data.frame(province2)
```
Clean region_1.
```{r}
region_1.2 <- stringi::stri_trans_general(WMDF$region_1, "latin-ascii")
```
New region_1.
```{r}
region_1.3 <- as.data.frame(region_1.2)
```
Clean variety.
```{r}
variety2 <- stringi::stri_trans_general(WMDF$variety, "latin-ascii")
```
New variety.
```{r}
variety3 <- as.data.frame(variety2)
```
Clean winery.
```{r}
winery2 <- stringi::stri_trans_general(WMDF$winery, "latin-ascii")
```
New winery.
```{r}
winery3 <- as.data.frame(winery2)
```
New WineMag dataframe.
```{r}
WineMag <- cbind(WMDF$country, description3, designation3, WMDF$points, WMDF$price, province3, region_1.3, WMDF$region_2, variety3, winery3)
```
Write CSV file.  Clean dataset now has 97,851 records with 10 attributes.
```{r}
write_csv(WineMag, path = "WineMag.csv")
```











